Nanophotonic Computational Design

### 2013-05-07 paper/paper.tex
Revision sent to Jelena.

\begin{thebibliography}{99}
\bibitem{lipson} 
    A. Gondarenko, and M. Lipson, 
    ``Low modal volume dipole-like dielectric slab resonator,'' 
    Opt. Express \textbf{16} 17689-17694 (2008).  
\bibitem{yablo} C.-Y. Kao, S. Osher, and E. Yablonovitch, 
    ``Maximizing band gaps in two-dimensional photonic crystals by using level set methods,'' Appl. Phys. B \textbf{81}, 235-244 (2005).
\bibitem{levi}
    P. Seliger, M. Mahvash, C. Wang, and A. Levi,
    ''Optimization of aperiodic dielectric structures,''
    J. Appl. Phys. \textbf{100}, 34310-6 (2006).
\bibitem{johnson} A. Oskooi, A. Mutapcic, S. Noda, J. D. Joannopoulos, S. P. Boyd, and S. G. Johnson, 
    ``Robust optimization of adiabatic tapers for coupling to slow-light photonic-crystal waveguides,'' Opt. Express \textbf{20}, 21558-21575 (2012).
\bibitem{sigmund1} Y. Elesin, B. S. Lazarov, J. S. Jensen, O. Sigmund, 
    ``Design of robust and efficient photonic switches using topology optimization,''
    Photonics and Nanostructures - Fundamentals and Applications \textbf{10}, 
    153–165 (2012).
\bibitem{jameson} L. Martinelli, and A. Jameson, 
    ``Computational Aerodynamics: Solvers and Shape Optimization,''
    J. Heat Transfer \textbf{135}, 011002 (2013).
\bibitem{bendsoe} M. P. Bendsoe, O. Sigmund,
    ``Material interpolation schemes in topology optimization,''
    Archive of Applied Mechanics \textbf{69}, 635-654 (1999).
\bibitem{miller} D. A. B. Miller, 
    ``All linear optical devices are mode converters,'' 
    Opt. Express \textbf{20}, 23985-23993 (2012).
\bibitem{ob1_wg} J. Lu, J. Vuckovic, ``Objective-first design of high-efficiency, small-footprint couplers between arbitrary nanophotonic waveguide modes,'' 
    Opt. Express \textbf{20}, 7221-7236 (2012)
\bibitem{admm} S. Boyd, N. Parikh, E. Chu, B. Peleato, and J. Eckstein,
    ``Distributed Optimization and Statistical Learning via the Alternating Direction Method of Multipliers,''
    Foundations and Trends in Machine Learning \textbf{3}, 1-122 (2011). 
\bibitem{wonseok} W. Shin and S. Fan, ``Choice of the perfectly matched layer boundary condition for frequency-domain Maxwell’s equations solvers,''
    J. of Comp. Phys \textbf{231}, 3406-3431 (2012).
\bibitem{levelset} S. Osher and R. Fedkiw, \emph{Level Set Methods and Dynamic Implicit Surfaces: 1st Edition} (Springer, 2002).
\bibitem{2dsplitter} Y. Jiao, S. Fan, and D. A. B. Miller, 
    ``Demonstrations of systematic photonic crystal design and optimization by low rank adjustment: an extremely compact mode separator'', Opt. Lett. \textbf{30}, 140-142 (2005).
\bibitem{mode1} J. Castro, D. F. Geraghty, S. Honkanen, C. M. Greiner, D. Iazikov, and T. W. Mossberg, ``Demonstration of mode conversion using anti-symmetric waveguide Bragg gratings,'' Opt. Express \textbf{13}, 4180-4184 (2005).
\bibitem{mode2} E. Khoo, A. Liu, and J. Wu, ``Nonuniform photonic crystal taper for high-efficiency mode coupling,'' Opt. Express \textbf{13}, 7748-7759 (2005).
\bibitem{mode3} P. Sanchis, J. Marti, J. Blasco, A. Martinez, and A. Garcia, ``Mode matching technique for highly efficient coupling between dielectric waveguides and planar photonic crystal circuits,'' Opt. Express \textbf{10}, 1391-1397 (2002). 
\bibitem{forchel} T. D. Happ, M. Kamp, and A. Forchel, 
    ``Photonic crystal tapers for ultracompact mode conversion,'' 
    Opt. Lett. \textbf{26}, 1102-1104 (2001) 
\bibitem{baets}F. Van Laere, G. Roelkens, M. Ayre, J. Schrauwen, D. Taillaert, D. Van Thourhout, T. F. Krauss, and R. Baets,
    ``Compact and highly efficient grating couplers between optical fiber and nanophotonic waveguides,''
    J. of Lightwave Tech. \textbf{25}, 151-156 (2007).
\bibitem{robust1} F. Wang, J. S. Jensen, and O. Sigmund, ``Robust topology optimization of photonic crystal waveguides with tailored dispersion properties,'' J. Opt. Soc. Am. B \textbf{28}, 387-397 (2011).
\end{thebibliography}

@@

Increasingly sophisticated optimization strategies for 
    specific nanophotonic devices have been developed;
    ranging from the use of genetic algorithms\cite{lipson},
    level-set methods\cite{yablo}, 
    tweaking the size and shape of specific canonical gemetric shapes\cite{levi,johnson},
    and even utilizing the full parameter space\cite{sigmund1}.

    \cite{jameson}
    \cite{bendsoe}

@@

For this reason, we have developed and implemented a computational method
    which is able to use the full parameter space 
    to design linear nanophotonic components in three dimensions.
Critically, our method requires no user intervention or manual tuning.
Instead, a \emph{design-by-specification} scheme is used 
    to produce designs based solely on a user's performance specification.

@@

Lastly, all our results are produced by simply specifying
    the functionality and performance of the desired device,
    which suggests that our method may indeed by able
    to design \emph{all} linear nanophotonic devices.

@@

@@

We show, through the design of mode converters for both the TE- and TM-polarized
    waveguide modes, 
    that our method is indeed fully three-dimensional.
Additionally, the devices also demonstrate very small device footprints 
    ($1.6 \times 2.4$ microns for this device in particular
    operating at 1550 nm wavelength).

@@

Although the performance specification (\fig{3Db_fib_te})
    desired a coupling efficiency above 90\%,
    only 51.5\% efficiency was achieved (\fig{fib_te});
    however, this likely remains the highest efficiency demonstrated
    in a compact fiber coupler.

@@

\subsubsection{Fabrication-robustness of broadband wavelength splitter}
A analysis with regard to fabrication-error
    was also performed on the broadband wavelength splitter.
The specific fabrication error was a general over- or under-etch
    of the device (input and output waveguides unaffected).
\Fig{analysis_fab_shift2} reveals that up to 8 nm of over- or under-etching
    can be sustained before performance falls below 70\%, at the central
    operating wavelengths.
The structural variations at 8 nm of etch error are shown in \fig{top_wdm_overunder}.

>>>>

\begin{thebibliography}{99}
\bibitem{lipson} 
    A. Gondarenko, and M. Lipson, 
    ``Low modal volume dipole-like dielectric slab resonator,'' 
    Opt. Express \textbf{16} 17689-17694 (2008).  
\bibitem{yablo} C.-Y. Kao, S. Osher, and E. Yablonovitch, 
    ``Maximizing band gaps in two-dimensional photonic crystals by using level set methods,'' Appl. Phys. B \textbf{81}, 235-244 (2005).
\bibitem{levi}
    P. Seliger, M. Mahvash, C. Wang, and A. Levi,
    ''Optimization of aperiodic dielectric structures,''
    J. Appl. Phys. \textbf{100}, 34310-6 (2006).
\bibitem{johnson} A. Oskooi, A. Mutapcic, S. Noda, J. D. Joannopoulos, S. P. Boyd, and S. G. Johnson, 
    ``Robust optimization of adiabatic tapers for coupling to slow-light photonic-crystal waveguides,'' Opt. Express \textbf{20}, 21558-21575 (2012).
\bibitem{sigmund1} Y. Elesin, B. S. Lazarov, J. S. Jensen, O. Sigmund, 
    ``Design of robust and efficient photonic switches using topology optimization,''
    Photonics and Nanostructures - Fundamentals and Applications \textbf{10}, 
    153–165 (2012).
\bibitem{jameson} L. Martinelli, and A. Jameson, 
    ``Computational Aerodynamics: Solvers and Shape Optimization,''
    J. Heat Transfer \textbf{135}, 011002 (2013).
\bibitem{bendsoe} M. P. Bendsoe, O. Sigmund,
    ``Material interpolation schemes in topology optimization,''
    Archive of Applied Mechanics \textbf{69}, 635-654 (1999).
\bibitem{ob1_wg} J. Lu, J. Vuckovic, ``Objective-first design of high-efficiency, small-footprint couplers between arbitrary nanophotonic waveguide modes,'' 
    Opt. Express \textbf{20}, 7221-7236 (2012)
\bibitem{miller} D. A. B. Miller, 
    ``All linear optical devices are mode converters,'' 
    Opt. Express \textbf{20}, 23985-23993 (2012).
\bibitem{admm} S. Boyd, N. Parikh, E. Chu, B. Peleato, and J. Eckstein,
    ``Distributed Optimization and Statistical Learning via the Alternating Direction Method of Multipliers,''
    Foundations and Trends in Machine Learning \textbf{3}, 1-122 (2011). 
\bibitem{wonseok} W. Shin and S. Fan, ``Choice of the perfectly matched layer boundary condition for frequency-domain Maxwell’s equations solvers,''
    J. of Comp. Phys \textbf{231}, 3406-3431 (2012).
\bibitem{levelset} S. Osher and R. Fedkiw, \emph{Level Set Methods and Dynamic Implicit Surfaces: 1st Edition} (Springer, 2002).
\bibitem{2dsplitter} Y. Jiao, S. Fan, and D. A. B. Miller, 
    ``Demonstrations of systematic photonic crystal design and optimization by low rank adjustment: an extremely compact mode separator'', Opt. Lett. \textbf{30}, 140-142 (2005).
\bibitem{mode1} J. Castro, D. F. Geraghty, S. Honkanen, C. M. Greiner, D. Iazikov, and T. W. Mossberg, ``Demonstration of mode conversion using anti-symmetric waveguide Bragg gratings,'' Opt. Express \textbf{13}, 4180-4184 (2005).
\bibitem{mode2} E. Khoo, A. Liu, and J. Wu, ``Nonuniform photonic crystal taper for high-efficiency mode coupling,'' Opt. Express \textbf{13}, 7748-7759 (2005).
\bibitem{mode3} P. Sanchis, J. Marti, J. Blasco, A. Martinez, and A. Garcia, ``Mode matching technique for highly efficient coupling between dielectric waveguides and planar photonic crystal circuits,'' Opt. Express \textbf{10}, 1391-1397 (2002). 
\bibitem{forchel} T. D. Happ, M. Kamp, and A. Forchel, 
    ``Photonic crystal tapers for ultracompact mode conversion,'' 
    Opt. Lett. \textbf{26}, 1102-1104 (2001) 
\bibitem{baets}F. Van Laere, G. Roelkens, M. Ayre, J. Schrauwen, D. Taillaert, D. Van Thourhout, T. F. Krauss, and R. Baets,
    ``Compact and highly efficient grating couplers between optical fiber and nanophotonic waveguides,''
    J. of Lightwave Tech. \textbf{25}, 151-156 (2007).
\bibitem{robust1} F. Wang, J. S. Jensen, and O. Sigmund, ``Robust topology optimization of photonic crystal waveguides with tailored dispersion properties,'' J. Opt. Soc. Am. B \textbf{28}, 387-397 (2011).
\end{thebibliography}

@@

@@

To this end, increasingly sophisticated optimization strategies for 
    specific nanophotonic devices have been developed
    which are able to search this parameter space.
These range from the use of genetic algorithms\cite{lipson},
    level-set methods\cite{yablo}, 
    and the optimization of specific geometric parameters\cite{levi,johnson}.
Additionally, much success has been attained\cite{sigmund1}
    through the use of topological design methods that
    have traditionally been applied in other fields\cite{jameson,bendsoe}.

In this work, we have endeavored to extend the current art
    by developing and implementing a general computational method
    for nanophotonic design which extends across all linear nanophotonic devices.
Specifically, our method uses the full parameter space 
    to design linear nanophotonic components in three dimensions.
Critically, our method requires no user intervention or manual tuning.
Instead, a \emph{design-by-specification} scheme is used 
    to produce designs based solely on a user's performance specification.

@@

Lastly, all our results are produced by simply specifying
    the functionality and performance of the desired linear device
    in terms of mode conversion\cite{miller}.
The variety of successful designs presented 
    suggests that our method may indeed by able
    to design \emph{all} linear nanophotonic devices.

@@

    We note that although the performance description of nearly all
        linear devices can be specified using such a constraint,
        the constraint itself is \emph{not} linear (neither is it convex).
    Other performance objectives are indeed possible which are nonlinear 
        and nonconvex (such as optimizing for energy density or Q-factor)---convex 
        objectives would be straightforward to implement,
        while some non-convex objectives would require clever heurestics.

@@

We show, through the design of mode converters for both the TE- and TM-polarized
    waveguide modes, 
    that our method is indeed fully three-dimensional.
Additionally, the devices also demonstrate very small device footprints 
    ($1.6 \times 2.4$ microns for this device in particular
    operating at 1550 nm wavelength).
Lastly, our devices convert between modes of opposing symmetry.

This is in contrast to prior art\cite{mode1,mode2,mode3,forchel},
    which is largely limited to two-dimensional designs,
    where mode-coupling is between same-symmetry modes,
    and devices are often larger in the transverse direction
    because of the need for photonic crystal mirrors.

@@

Although the performance specification (\fig{3Db_fib_te})
    desired a coupling efficiency above 90\%,
    only 51.5\% efficiency was achieved (\fig{fib_te}).

@@

\subsubsection{Fabrication-robustness of broadband wavelength splitter}
A analysis with regard to fabrication-error
    was also performed on the broadband wavelength splitter.
The specific fabrication error was a general over- or under-etch
    of the device (input and output waveguides unaffected)\cite{robust1}.
\Fig{analysis_fab_shift2} reveals that up to 8 nm of over- or under-etching
    can be sustained before performance falls below 70\%, at the central
    operating wavelengths.
The structural variations at 8 nm of etch error are shown in \fig{top_wdm_overunder}.
